---
title: "Class 11: Candy Project"
author: "Jonathan Cifuentes A17238145"
format: html
---

In today's class we will examine 538 candy data and see if this helps us gain some more feeling for how PCA and other methods work.

```{r}
candy <- read.csv("candy-data.csv", row.names =1)
head(candy)
```

```{r}
row.names(candy)
```

> Q1. How many different candy types are in this dataset?

```{r}
nrow(candy)
```

> Q2. How many fruity candy types are in the dataset?

```{r}
table(candy$fruity)
```
```{r}
sum(candy$fruity)
```

> What are these fruity candy?

We can use the ==
```{r}
#== 1 calls forth the candies that were fruity, brackets select that and row.names brings the name of those candies
row.names(candy[candy$fruity== 1, ])
```
## How often does my favorite candy win?

```{r}
candy["Twix", "winpercent"]
candy["Skittles original", "winpercent"]
```

> Q4. What is the winpercetn value for "kit kat"?

```{r}
candy["Kit Kat", "winpercent"]
```
> and toosie rolls snack bars?

```{r}
candy["Tootsie Roll Snack Bars", "winpercent"]
```

```{r}
library("skimr")
skim(candy)
```

> Q6. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?

Yeap, the `winpercent` column is on a 0:100 scale and all others appear to be 0:1 scale
`

> Q7. What do you think a zero and one represent for the candy$chocolate column?
 
Whether or not the candy contains chocolate. yes/no question

> Q8. Plot a histogram of winpercent values

```{r}
library(ggplot2)
ggplot(candy, aes(winpercent)) +geom_histogram(bins=15
                                               )
```

> Q10 Is the distribution of winpercent values symmetrical?
 
 no
> Q11. I sthe center of the distribution above or below 50%?

Below 50% with a mean of:
```{r}
# mean(candy$winpercent)
candy$winpercent[]
```

> Q11. On average is chocolate candy higher or lower ranked than fruit candy?

```{r}
#harder route
mean(candy$winpercent[which(candy$chocolate==1)])

mean(candy$winpercent[which(candy$fruity==1)])
```

```{r}
#easier rout
#filter/select/subset to just chocolate rows
chococandy <- candy[as.logical(candy$chocolate), ]
# Get their winpercent values
chocowinpercent <-  chococandy$winpercent
#calculate their mean winpercent value
mean(chocowinpercent)
```
```{r}
#easier rout
#filter/select/subset to just chocolate rows
fruitycandy <- candy[as.logical(candy$fruity), ]
# Get their winpercent values
fruitywinpercent <-  fruitycandy$winpercent
#calculate their mean winpercent value
mean(fruitywinpercent)
```
##Overal Candy Rankings

There is a base R function called `sort(` for, guess what sorting vectors input.

```{r}
x <- c(5, 2, 10)

#sort(x, decreasing=T)
sort(x)
```
The buddy function to `sort()` that is often more useful is called `order()`. It returns the "indices" of the input that would result in it being sorted.
```{r}
order(x)
x[order(x)]
```
> Q13. What are the five least liked candy types in this set?

I can order by `winpercent`
```{r}
ord <- order(candy$winpercent)
head(candy[ord,], 5)

```

> Q14. What are the top 5 all time favorite candy types out of this set?

```{r}
ord <- order(candy$winpercent, decreasing=T)
head(candy[ord,], 5)
```

```{r}
candybar <- ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy), winpercent)) +
  geom_col()
```

```{r}
my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "chocolate"
my_cols[as.logical(candy$bar)] = "brown"
my_cols[as.logical(candy$fruity)] = "pink"
```

```{r}
candybar+ geom_col(fill=my_cols)
```
> Q17. What is the worst ranked chocolate candy?

sixlets
> Q18. What is the best ranked fruity candy?

starburst

> Whats the best candy for the least money?

```{r}
library(ggrepel)
ggplot(candy) +aes(winpercent, pricepercent, label= rownames(candy))+ geom_point(col=my_cols) + geom_text()+ geom_text_repel(col=my_cols, size=3.3, max.overlaps = 8)
```

# 5 Exploring the coloration structure

Pearson correlation goes between -1 and +1 with zero indicatin gno correlation, and vlues close to one being highly (ani) correlated.
```{r}
library(corrplot)
cij <- cor(candy)
corrplot(cij)
```

> Q22. Examining this plot what two variables are anti-correlated (i.e. have minus values)?

Fruit/chocolate & bar/pluribus
> Q23. Similarly, what two variables are most positively correlated?

winpercent/chocolate

# 6 Principal Component Analysis

The base R function for PCA is called `prcomp()` and we can set the "scale=TRUE/FALSE".

```{r}
pca <- prcomp(candy, scale=T)
summary(pca)
```

The main result of PCA- IE the new PC plot ( projection of candy on our new PC axis) is contained in `pca$x`

```{r}
#need to convert as data frame because ggplot can only read the prcomp as that
pc<- as.data.frame(pca$x)
ggplot(pc)+ aes(PC1, PC2, label=rownames(pc)) + geom_point(col=my_cols)+ geom_text_repel(max.overlaps = 5)
```

>Q24. What original variables are picked up strongly by PC1 in the positive direction? Do these make sense to you?

```{r}
par(mar=c(8,4,2,2))
barplot(pca$rotation[,1], las=2, ylab="PC1 Contribution")
```
Fruity, hard, pluribus all had positive inflections for PC1
