---
title: "Class 09 Structural Bioinformatics 1"
format: html
---

#Comparatove analysis of ADK
ADK (Adenelate Kinase) is a n important drug target and we would love to know how it works - ie molecular mechanism.]

There has been lots of work done on this protein due to it's importance including lots of crystal structure.

```{r}
library(bio3d)
```

```{r}
aa <- get.seq("1ake_A")
aa
```

We will begin by 

We can now run BLAST with this sequence
```{r}
b <- blast.pdb(aa)
```

```{r}
hits <- plot(b)
```

Let's see whats it our 'hits'
```{r}
hits$pdb.id
```
Now we can download all these PDB structure files:
```{r}
files <- get.pdb(hits$pdb.id, path="pdbs", split=TRUE, gzip=TRUE)
```
Now i want to align and superimpose these structures which are all over the place.
```{r}
# Align releated PDBs
pdbs <- pdbaln(files, fit = TRUE, exefile="msa")
```

```{r}
# Vector containing PDB codes for figure axis
ids <- basename.pdb(pdbs$id)

# Draw schematic alignment
plot(pdbs, labels=ids)
```

```{r}
#annotate
anno <- pdb.annotate(ids)
unique(anno$source)
```
```{r}
anno
```

```{r}
# Perform PCA
pc.xray <- pca(pdbs)
plot(pc.xray)
```

```{r}
# Calculate RMSD
rd <- rmsd(pdbs)

# Structure-based clustering
hc.rd <- hclust(dist(rd))
grps.rd <- cutree(hc.rd, k=3)

plot(pc.xray, 1:2, col="grey50", bg=grps.rd, pch=21, cex=1)
```

We can make a wee movie - also called a trajectory of the major differences (ie structural displacements) of ADK.
```{r}
#Visualize first principal component makes a file of the protein in mol view
pc1 <- mktrj(pc.xray, pc=1, file="pc_1.pdb")
```

> Q10. Which of the packages above is found only on BioConductor and not CRAN? 

MSA
> Q11. Which of the above packages is not found on BioConductor or CRAN?: 

Bio3D-view
> Q12. True or False? Functions from the devtools package can be used to install packages from GitHub and BitBucket? 

true
